Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-08 mr-review.Rmd
merge fixes

In the case of continuous exposures and outcomes, researchers have taken several approaches to extend the methods and models for 
single-SNP MR to accommodate multiple SNPs. For two stage analyses, one may perform 
a sequence of two linear regressions, as in the single-SNP MR models (Equations \@ref(eq:lm3) and \@ref(eq:lm4)). 
For an analysis
with $S$ SNP instruments, one performs two stage analysis by first regressing the 
exposure $n$ vector $X$ on the $n$ by $S$ matrix of SNP instruments. Taking the fitted exposure values 
from this regression, the second regression involves regressing the outcome variable $Y$ on 
the fitted values.

@@

Early MR analyses avoided confounding from horizontal pleiotropy by discarding 
instrumental SNPs that might be associated with the outcome [@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-1;@zeng_causal_2019-2].
More recent methods have attempted to model horizontal pleiotropy [@morrison_mendelian_2020;@yuan_testing_2020;@yuan_likelihood_2021]. CAUSE [@morrison_mendelian_2020] mdoels two varieties of horizontal pleiotropy:

@@

>>>>

In the case of continuous exposures and outcomes, researchers have taken several approaches to extend the methods and models for 
single-SNP MR to accommodate multiple SNPs. For two stage analyses, one may perform 
a sequence of two linear regressions, as in the single-SNP MR models (Equations \@ref(eq:lm3) and \@ref(eq:lm4)). 
For an analysis
with $S$ SNP instruments, one performs two stage analysis by first regressing the 
exposure $n$ vector $X$ on the $n$ by $S$ matrix of SNP instruments. Taking the fitted exposure values 
from this regression, the second stage involves regressing the outcome variable $Y$ on 
the fitted values, $\hat X$.

@@

Researchers have recognized two types of horizontal pleiotropy [@yuan_likelihood_2021]. The first occurs via 
exposure-independent paths. The resulting horizontal pleiotropic effects are independent of the
SNP-exposure relationships. The second type of horizontal pleiotropy manifests when 
in the presence of unobserved exposure-outcome confounding. It induces correlation between 
horizontal pleiotropic effects and SNP-exposure effects. Both types of horizontal pleiotropy 
violate standard MR modeling assumptions and can bias causal effect estimates and can increase false 
discoveries [@yuan_likelihood_2021].
Early MR analyses avoided confounding from horizontal pleiotropy by discarding 
instrumental SNPs that might be associated with the outcome [@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-1;@zeng_causal_2019-2].
More recent methods have attempted to model horizontal pleiotropy [@morrison_mendelian_2020;@yuan_testing_2020;@yuan_likelihood_2021]. CAUSE [@morrison_mendelian_2020] mdoels two varieties of horizontal pleiotropy:

@@

Early MR methods account manage horizontal pleiotropy by removing instrumental SNPs 
that may be associated with the outcome variable [@verbanck_detection_2018;@zhu_causal_2018;@zeng_causal_2019-1;@zeng_causal_2019-2].
